Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering

### 2016-07-12 paper/structureScene.tex
moving dataset to experiments

The proposed approach is motivated in Section \ref{sec:motivations} and a brief review the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}. We describe the scattering transform in Section \ref{sec:scattering}, discuss post-processing transformations in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}. In Section \ref{sec:experiments} several experiments are proposed for an unsupervised setting, \ie acoustic scene similarity retrieval (ASSR) and for a supervised setting, \ie acoustic scene classification (ASC). Results are subsequently reported in Section \ref{sec:results}.

\section{Motivations} \label{sec:motivations}

@@

Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks. \ja{Since we go into so much detail on ASC, shouldn't we at lest describe what these works do? Some of this is already described in the motivations section, so perhaps we should move some over in here?}
An overview of the efforts for the latter task is provided by the DCASE2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare differrent approaches to scene classification. Table \ref{tab:dcase} summarizes the results of the submitted algorithms where most \ja{?} of scene predictions are achieved by majority voting of predictions over features. \ja{What about the other algorithms? Why are we excluding them here?}

\begin{table}
\begin{center}
\begin{tabular}{lcl}
& Acc. \%  &  \multicolumn{1}{c}{Design}  \\
\cite{olivetti2013wonder} OE & 17 & compression distance, random forests \\
\cite{elizalde2013vector} ELF & 55 & i-vector, LDA \\
\cite{krijnders2013tone} KH & 55 & cochleogram, SVMs \\
\cite{patil2013multiresolution} PE & 58 & 2D Gabor filters, SVMs \\
\cite{lee2013acoustic} NHL & 60 & UL of mel spectra, SVMs \\
\cite{nogueira2013sound} NR1 & 60 & MFCCs and ILD features, SVMs \\
\cite{chum2013ieee} CHR & 65 & sparsity features, HMMs  \\
\cite{geiger2013large} GSR & 69 & mel \ja{?} and MFCCs, SVMs \\
\cite{rakotomamonjy2015histogram} RG & 69 & histogram of gradient (HOG), SVMs \\
\cite{li2013auditory} LTT & 72 & MFCCs, tree baggers \\
\cite{roma2013recurrence} RNH & 76 & RQA features stacked with MFCCs, SVMs \\
\end{tabular}
\end{center}
\caption{Designs submitted to the DCASE2013 ASC challenge. RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning. \label{tab:dcase}}
\end{table}

In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task. \ja{It's not clear to me how the first part implies the second part here} Similarly, we have picked linear SVMs as our classifier for the ASC task, since these are a standard tool in the community and allows for us to focus our analysis on the earlier processing stages. Moreover, since MFCCs are well-established features which perform well for these task, we shall include them baseline features in all experiments described in this paper.

@@

For most of the experiments reported in this paper, we will consider the DCASE2013 dataset \cite{giannoulis2013database, 7100934}. To build this dataset, three different recordists visited a wide variety of locations in Greater London over
a period of several months and in each
scene recorded a few minutes of audio. No
systematic variations in the recordings covaried with scene
type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a fivefold cross validation scheme. The folds used in this paper are the same as the ones used during the challenge. \ja{Why are we describing the dataset here in the background section? Wouldn't it be more appropriate under experiments?}

There are also two larger datasets: the
Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge \cite{Mesaros2016_EUSIPCO}. The latter is also considered in this paper as a validation of the proposed approach. \ja{What does this last sentence mean?}

In this paper, we shall demonstrate that the use of features based on the wavelet scattering transform together with object-based representation of the acoustic scene allows us to achieve state-of-the-art \ja{?} performance in both the ASSR and ASC paradigms.

@@

The wavelet transform of an audio signal
$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets.
Applying pointwise complex modulus the transform yields
the wavelet scalogram
\begin{equation}
\boldsymbol{x_1}(t, \gamma_1)
= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t)\mbox{.}
\end{equation}
The wavelet scalogram bears resemblance to the constant-Q transform (CQT),
which is derived from the short-term Fourier transform (STFT) by averaging the frequency
axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q_1}\xi_1$.
Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
However, contrary to the CQT, the wavelet scalogram reaches the Heisenberg
theoretical limit of optimal time-frequency localization across the whole
frequency range, whereas the temporal resolution of the CQT, at least in its original implementation \cite{Brown1992}, is fixed by the support of the STFT analyzing window.
Therefore, the wavelet scalogram has a better temporal localization at high
frequencies than the CQT, at the expense of a greater computational cost.
This increase in time-frequency localization comes at the expense of a greater computational cost, the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
Yet, it allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.

@@

Taking the logarithm of a magnitude spectrum is ubiquitous in audio signal processing.
Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. 
We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone, a source of spurious variability in scene classification.
Logarithmic compression can linearize this dependency, facilitating the construction of a powerful invariant at the classifier stage.

@@

@@

\subsubsection*{Early \vs late integration}
late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
This means that, despite its discriminative power at fine scales, the scattering transform is unable to segregate distinct acoustic events beyond the scale of about one second.

@@

\subsubsection*{Participation to the DCASE 2016 challenge}
In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge, whose results are forthcoming.
This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel.
Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.

Our results on the development dataset are reported in Table \ref{table:dcase2016}, in comparison with a baseline of $MFCCs$ coefficients and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
The average accuracy across classes is brought from $70.8\%$ to $79.4\%$.
The results on the evaluation dataset are awaiting publication by the DCASE organizing committee, and will be disclosed in August 2016.

>>>>

Motivations of the proposed approach and a brief review of the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}. We describe the scattering transform in Section \ref{sec:scattering}, discuss post-processing transformations in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}. In Section \ref{sec:experiments} several experiments are proposed for an unsupervised setting, \ie acoustic scene similarity retrieval (ASSR) and for a supervised setting, \ie acoustic scene classification (ASC). Results are subsequently reported in Section \ref{sec:results}.

@@

BOF approaches treats each observations equally, whereas the above cited results show that only a few events scarcely occurring may have be more important than events or texture present throughout the whole scene duration, thus advocating for a more selective approach such as the one presented in Section \ref{sec:object} for the ASSR task.

%\section{Background} \label{sec:soa}

Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks. \ja{Since we go into so much detail on ASC, shouldn't we at lest describe what these works do? Some of this is already described in the motivations section, so perhaps we should move some over in here?}
An overview of the efforts for the latter task is provided by the DCASE2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification. Table \ref{tab:dcase} summarizes the results of the submitted algorithms where most \ja{?} of scene predictions are achieved by majority voting of predictions over features. \ja{What about the other algorithms? Why are we excluding them here?}

\begin{table}
\begin{center}
\begin{tabular}{lcl}
& Acc. \%  &  \multicolumn{1}{c}{Design}  \\
\cite{olivetti2013wonder} OE & 17 & compression distance, random forests \\
\cite{elizalde2013vector} ELF & 55 & i-vector, LDA \\
\cite{krijnders2013tone} KH & 55 & cochleogram, SVMs \\
\cite{patil2013multiresolution} PE & 58 & 2D Gabor filters, SVMs \\
\cite{lee2013acoustic} NHL & 60 & UL of mel spectra, SVMs \\
\cite{nogueira2013sound} NR1 & 60 & MFCCs and ILD features, SVMs \\
\cite{chum2013ieee} CHR & 65 & sparsity features, HMMs  \\
\cite{geiger2013large} GSR & 69 & Mel-frequency Spectra and MFCCs, SVMs \\
\cite{rakotomamonjy2015histogram} RG & 69 & histogram of gradient (HOG), SVMs \\
\cite{li2013auditory} LTT & 72 & MFCCs, tree baggers \\
\cite{roma2013recurrence} RNH & 76 & RQA features stacked with MFCCs, SVMs \\
\end{tabular}
\end{center}
\caption{Designs submitted to the DCASE2013 ASC challenge. RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning. \label{tab:dcase}}
\end{table}

In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task. \ja{It's not clear to me how the first part implies the second part here} Similarly, we have picked SVMs as our classifier for the ASC task, since these are a standard tool in the community and allows for us to focus our analysis on the earlier processing stages. Moreover, since MFCCs are well-established features which perform well for these task, we shall include them baseline features in all experiments described in this paper.

@@

 % The latter is also considered in this paper as a validation of the proposed approach. \ja{What does this last sentence mean?}

%In this paper, we shall demonstrate that the use of features based on the wavelet scattering transform together with object-based representation of the acoustic scene allows us to achieve state-of-the-art \ja{?} performance in both the ASSR and ASC paradigms.

@@

The wavelet transform of an audio signal
$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets.
Applying pointwise complex modulus the transform yields
the wavelet scalogram
\begin{equation}
\boldsymbol{x_1}(t, \gamma_1)
= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t)\mbox{.}
\end{equation}
The wavelet scalogram bears resemblance to the constant-Q transform (CQT),
which is derived from the short-term Fourier transform (STFT) by averaging the frequency
axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q_1}\xi_1$.
Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
However, contrary to the CQT, the wavelet scalogram reaches the Heisenberg
theoretical limit of optimal time-frequency localization across the whole
frequency range, whereas the temporal resolution of the CQT is fixed by the support of the STFT analyzing window. \ja{This is not true. See for example Sch\"orkhuber and Klapuri 2010.}
Therefore, the wavelet scalogram has a better temporal localization at high
frequencies than the CQT, at the expense of a greater computational cost. \ja{Why is the cost greater?}
This property allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.

@@

Taking the logarithm of a magnitude spectrum is ubiquitous in audio signal processing.
Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. \ja{Cite?}
We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone, a source of spurious variability in scene classification.
Logarithmic compression can linearize this dependency, facilitating the construction of a powerful invariant at the classifier stage.

@@

\subsection{Datasets}

%For most of the experiments reported in this paper, we will consider the DCASE2013 dataset \cite{giannoulis2013database, 7100934}. Larger datasets exists: the
%Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge.%

In the experiments reported in this paper, the DCASE2013 \cite{giannoulis2013database, 7100934} and the public part of the DCASE2016 datasets  \cite{Mesaros2016_EUSIPCO} are considered. To build the DCASE2013 dataset, three different recordists visited a wide variety of locations in Greater London over
a period of several months and in each
scene recorded a few minutes of audio. No
systematic variations in the recordings covaried with scene
type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a fivefold cross validation scheme. The folds used in this paper are the same as the ones used during the challenge.

@@

\subsubsection*{Early \vs late integration}
late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.

@@

\subsubsection*{Participation to the DCASE 2016 challenge}
In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge.

Our results on the development dataset are reported in Table \ref{table:dcase2016}, in comparison with a baseline of $MFCCs$ coefficients and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
The average accuracy across classes is brought from $70.8\%$ to $79.4\%$.
It appears that log-scattering features compare favorably to 
The results on the evaluation dataset are awaiting publication by the DCASE organizing committee, and will be disclosed in August 2016.
